Watch Out for the Bully! Job Interference Study on Dragonfly Network

### 2016-04-09 background.tex

\subsection{Dragonfly Network}
\label{sec:network}
The dragonfly is a two-level hierarchical topology, consisting of several groups connected by all-to-all links. In each group, there can be $a$ routers connected via all-to-all local channels. For each router, there can be $p$ compute nodes attached to it via terminal links. The router has $h$ global channels used as intergroup connections through which it connected to other routers in different groups. Thus, the radix of each router is $k = a+h+p-1$. Different computing centers could choose different values for  $a$, $h$, $p$ when deploying their dragonfly network. The adoption of proper $a$, $h$, $p$ involves many factors such as system scale, building cost and workload characteristics.

It is recommended that for the load balance purpose, a proper dragonfly configuration should follow $a=2p=2h$~\cite{kim-micro}. According to this configuration, the total number of groups, denoted as $g$ in the dragonfly network would be $g = a*h+1 $, the total number of compute nodes denoted as $N$ in the network would be $N = p*a*g $. An example dragonfly network is illustrated in Figure~\ref{fig:dragonfly-overview}. There are six routers in each group $(a=6)$, three compute nodes per router $(p=3)$, and three global channels per router $(h=3)$. This dragonfly network consists of 19 groups and 342 nodes in total.

@@

The routing policy refers to the strategy adopted to route  packets from the source router to the destination router. Previously studied routing policies for dragonfly network include minimal routing, adaptive routing~\cite{dally-dragonfly}, progressive adaptive routing~\cite{jiang} and some of their variations~\cite{won-prog-adaptive}. In this work we study three alternative routing policies considered by the community for dragonfly networks.

\textbf{Minimal:} In this policy, a packet takes the minimal (shortest) path from the source to the destination. The packet first route locally from the source node to the global channel leading to the destination group. It traverses the global channel to the destination group and routes locally to the destination node. 
Minimal routing can guarantee the minimum hops a packet takes from the source to the destination. However, it usually results in congestion along the minimal paths.

\textbf{Adaptive:} In this policy, the path a packet takes will be adaptively chosen between minimal and non-minimal paths, depending on the congestion situation along those paths. For the non-minimal path, an intermediate router in a separate group will be randomly chosen. The packet takes the intermediate router as a via point, connecting the source and destination router through two separate minimal paths. Adaptive routing can avoid hot-spots in the presence of congestion and collapse to minimal routing otherwise.

\textbf{Progressive Adaptive:} As opposed to adaptive routing, the decision to route a packet minimally is re-evaluated until a non-minimal route is chosen for the packet or as long as the packet is in its source group; the re-evaulation does not occur in intermediate groups~\cite{jiang}.
Progressive adaptive routing is capable of handling scenarios where the minimal route is congested but the source router has not been informed yet.

@@

For a parallel application requiring multiple compute nodes, job placement policy refers to the way of assigning the required number of nodes to the application by a system software such as batch scheduler~\cite{xu-cluster14}. In this work, we study two alternative placement policies considered by the community for dragonfly systems:

\textbf{Random Placement:} In this policy, an application gets the required number of nodes randomly selected from all the available nodes in the system. As illustrated in Figure~\ref{fig:dragonfly-overview}, $J1$ gets random allocation in which nodes are attached to different routers in different groups. Routers may be shared by different applications and more routers are involved in serving each application when random placement is in use. Random placement can distribute the tasks of an application uniformly across the network to avoid the possible local congestion.

\textbf{Contiguous Placement:} In this policy, the compute nodes are assigned to an application in a consecutive manner. The assignment first fills up a group, then crosses group boundaries and starts to fill the next group if necessary. As illustrated in Figure~\ref{fig:dragonfly-overview}, $J2$ gets eight nodes by contiguous placement. Contiguous placement confines the tasks of an application into the same group and uses minimum number of routers serving each application, which may result in local network congestion and increase the possibility of hot-spots.

>>>>

\subsection{Dragonfly Network}
\label{sec:network}
The dragonfly is a two-level hierarchical topology, consisting of several groups connected by all-to-all links. Each group consists of $a$ routers connected via all-to-all local channels. For each router, $p$ compute nodes are attached to it via terminal links, while $h$ links are used as global channels for intergroup connections. The resulting radix of each router is $k = a+h+p-1$. Different computing centers could choose different values for  $a$, $h$, $p$ when deploying their dragonfly network. The adoption of proper $a$, $h$, $p$ involves many factors such as system scale, building cost and workload characteristics.

It is recommended that for load balancing purposes, a proper dragonfly configuration should follow $a=2p=2h$~\cite{kim-micro}. According to this configuration, the total number of groups, denoted as $g$ in the dragonfly network would be $g = a*h+1 $, the total number of compute nodes denoted as $N$ in the network would be $N = p*a*g $. An example dragonfly network is illustrated in Figure~\ref{fig:dragonfly-overview}. There are six routers in each group $(a=6)$, three compute nodes per router $(p=3)$, and three global channels per router $(h=3)$. This dragonfly network consists of 19 groups and 342 nodes in total.

@@

The routing policy refers to the strategy adopted to route  packets from the source router to the destination router. Previously studied routing policies for dragonfly networks include minimal routing, adaptive routing~\cite{dally-dragonfly}, progressive adaptive routing~\cite{jiang} and variations thereof~\cite{won-prog-adaptive}. In this work we study three alternative routing policies considered by the community for dragonfly networks.

\textbf{Minimal:} In this policy, a packet takes the minimal (shortest) path from the source to the destination. The packet first routes locally from the source node to the global channel leading to the destination group. It traverses the global channel to the destination group and routes locally to the destination node. 
Minimal routing can guarantee the minimum hops a packet takes from the source to the destination. However, it usually results in congestion along the minimal paths.

\textbf{Adaptive:} In this policy, the path a packet \TODO{packet granularity or message granularity?} takes will be adaptively chosen between minimal and non-minimal paths, depending on the congestion situation along those paths. For the non-minimal path, an intermediate router in a separate group will be randomly chosen. The packet is forwarded to the intermediate router, connecting the source and destination groups through two separate minimal paths. Adaptive routing can avoid hot-spots in the presence of congestion and collapses to minimal routing otherwise.

\textbf{Progressive Adaptive:} As opposed to adaptive routing, the decision to adaptively route message packets is continually re-evaluated within the source group until a non-minimal route is chosen; the re-evaluation does not occur in intermediate groups~\cite{jiang}.
Progressive adaptive routing is capable of handling scenarios where the minimal route is congested but the source router has not been informed yet.

@@

For a parallel application requiring multiple compute nodes, the job placement policy refers to the way of assigning the required number of nodes to the application by system software such as the batch scheduler~\cite{xu-cluster14}. In this work, we study two alternative placement policies considered by the community for dragonfly systems:

\textbf{Random Placement:} In this policy, an application gets the required number of nodes randomly from the available nodes in the system. As illustrated in Figure~\ref{fig:dragonfly-overview}, $J1$ is randomly allocated nodes attached to different routers in different groups. Routers may be shared by different applications and more routers are involved in serving each application when random placement is in use. Random placement can distribute the tasks of an application uniformly across the network to avoid the possible local congestion.

\textbf{Contiguous Placement:} In this policy, the compute nodes are assigned to an application consecutively. The assignment first fills up a group, then crosses group boundaries as necessary. As illustrated in Figure~\ref{fig:dragonfly-overview}, $J2$ is allocated eight nodes by contiguous placement. Contiguous placement confines the tasks of an application into the same group and uses the minimum number of routers to serve each application, which may result in local network congestion and increase the possibility of hot-spots.

### 2016-04-09 methodology.tex

We utilize the CODES simulation toolkit (Co-Design of Multilayer Exascale Storage Architectures)~\cite{misbah-tpds}, which builds upon the ROSS parallel discrete event simulator~\cite{carothers_ross:_2002,barnes_warp_2013} to enable exploratory study of large scale systems of interest to the HPC community. CODES supports dragonfly~\cite{codes-dragonfly, misbah-tpds}, torus~\cite{misbah-pads-2014, ning-pads-2011}, and Slim Fly~\cite{wolfe-pads-2016} networks with flit-level high-fidelity simulation. It can drive these models through an MPI simulation layer utilizing traces generated by the DUMPI MPI trace library available as part of the SST macro toolkit~\cite{sst}. 
The behavior and performance of CODES dragonfly network model has been validated by Mubarak et al.\ \cite{codes-dragonfly} against booksim, a serial cycle-accurate interconnection network simulator~\cite{booksim-simulator}.

@@

We use a trace-driven approach to workload generation, choosing in particular three parallel application traces gathered to represent exascale workload behavior as part of the DOE Design Forward Project~\cite{designforward-webpage,designforward-traces} Specifically, we study communication traces representing the \emph{Algebraic MultiGrid Solver} (AMG), \emph{Geometric Multigrid V-Cycle from Production Elliptic Solver} (MultiGrid) and \emph{Crystal Router MiniApp} (CrystalRouter).

@@

The parameters for building the dragonfly network studied in our work are informed by the model proposed by Kim et al.~\cite{kim-micro}. 
In our dragonfly network, each group consists of $a = 8$ routers connected via all-to-all local channels. For each router, there could be $p = 4$ compute nodes attached to it via terminal links. The router has $h = 4$ global channels used as intergroup connections through which it connected to other routers in different groups. The radix of each router is $k = a+h+p-1 = 15$. 
The total number of groups is $g = a*h+1 = 33 $ and the total number of compute nodes is $N = p*a*g = 1056$. 
%\NOTE{What's the resulting router radix? What's the classic $a,h,p$ (or whatever they are :) parameters? done! } 
The dragonfly link bandwidths are 5.25 GiB/s for the local and terminal router-ports and 4.7 GiB/s for the global ports, inspired by the Cray Cascade system~\cite{faanes} 
%\NOTE{What are these bandwidths? done!}. 
In this work, we simulate the network performance and job interference across six different job placement and routing policy combinations, which are summarized in Table~\ref{tab: placement routing configs}.

@@

We analyze both the overall network performance and the performance of every single application.
Our analysis focuses on the following metrics:
\begin{itemize}

>>>>

We utilize the CODES simulation toolkit (Co-Design of Multilayer Exascale Storage Architectures)~\cite{misbah-tpds}, which builds upon the ROSS parallel discrete event simulator~\cite{carothers_ross:_2002,barnes_warp_2013} to enable exploratory study of large scale systems of interest to the HPC community. CODES supports dragonfly~\cite{codes-dragonfly, misbah-tpds}, torus~\cite{misbah-pads-2014, ning-pads-2011}, and Slim Fly~\cite{wolfe-pads-2016} networks with flit-level high-fidelity simulation. It can drive these models through an MPI simulation layer utilizing traces generated by the DUMPI MPI trace library available as part of the SST macro toolkit~\cite{sst}. 
The behavior and performance of the CODES dragonfly network model has been validated by Mubarak et al.\ \cite{codes-dragonfly} against BookSim, a serial cycle-accurate interconnection network simulator~\cite{booksim-simulator}.

@@

We use a trace-driven approach to workload generation, choosing in particular three parallel application traces gathered to represent exascale workload behavior as part of the DOE Design Forward Project~\cite{designforward-webpage,designforward-traces}. Specifically, we study communication traces representing the \emph{Algebraic MultiGrid Solver} (AMG), \emph{Geometric Multigrid V-Cycle from Production Elliptic Solver} (MultiGrid) and \emph{Crystal Router MiniApp} (CrystalRouter).

@@

The parameters for building the dragonfly network studied in our work are informed by the model proposed by Kim et al.~\cite{kim-micro}. 
In our dragonfly network, each group consists of $a = 8$ routers connected via all-to-all local channels. For each router, there are $p = 4$ compute nodes attached to it via terminal links. Each router also has $h = 4$ global channels used for intergroup connections. The radix of each router is hence $k = a+h+p-1 = 15$. 
The total number of groups is $g = a*h+1 = 33 $ and the total number of compute nodes is $N = p*a*g = 1056$. 
%\NOTE{What's the resulting router radix? What's the classic $a,h,p$ (or whatever they are :) parameters? done! } 
The dragonfly link bandwidths are 5.25 GiB/s for the local and terminal router-ports and 4.7 GiB/s for the global ports, inspired by the Cray Cascade system~\cite{faanes} 
%\NOTE{What are these bandwidths? done!}. 
In this work, we simulate the network performance and job interference across six different job placement and routing policy combinations, which are summarized in Table~\ref{tab: placement routing configs}.

@@

We analyze both the overall network performance and the performance of each application.
Our analysis focuses on the following metrics:
\begin{itemize}
